---
title: "Code Appendix"
author: "Laura Cosgrove"
date: "12/4/2019"
output: pdf_document
toc: true
---

```{r}
library(tidyverse)
library(geepack)
```

# Read data

```{r}
alc_dat <- readxl::read_xls("./data/ALCDEP.xls")

alc_dat <- janitor::clean_names(alc_dat) %>% 
  mutate(treatment = factor(treatment)) %>% 
  mutate(gender = factor(gender, levels = c(0,1), labels = c("Male", "Female")))

alc_dat_long <- alc_dat %>% 
  pivot_longer(nd0:nd60, names_to = "days", values_to = "drinks", names_prefix = "nd")
```

# Explore outcome distribution

```{r}
summary(alc_dat_long$drinks)

alc_dat_long %>% 
  group_by(gender, days, treatment) %>% 
  summarize(mean = mean(drinks),
            var = var(drinks),
            scale = var/mean)  %>% 
  ggplot(aes(x = mean, y = var, color = scale)) + geom_point() + geom_abline(intercept = 0, slope = 1) + geom_smooth(method = "lm", color = "grey", se = FALSE) + ggthemes::theme_few() + scale_x_continuous(limits = c(0, 200)) + scale_y_continuous(limits = c(0, 600)) + labs(title = "Potential overdispersion in Poisson outcome")
#still some overdispersion if poisson in the gender x treatment x time model; maybe need negative binomial model, maybe not

#normal approximation?
alc_dat_long %>% 
ggplot(aes(x = drinks)) + geom_histogram(bins = 60) + ggthemes::theme_few() + facet_grid(c("gender"))
#clearly not a normal distribution, clearly poisson

```

Choose a poisson-distributed outcome and check diagnostics for overdispersion.

# Exploratory Mean and Trajectory Plots

```{r}
alc_dat_long %>% 
  group_by(treatment, days) %>% 
  summarize(mean_drinks = mean(drinks),
            upper = mean_drinks + sd(drinks),
            lower = mean_drinks - sd(drinks)) %>% 
  ggplot(aes(x = days, color = treatment)) +
  geom_line(aes(y = mean_drinks, group = treatment), size = 1) + 
  geom_errorbar(aes(ymin = lower, ymax = upper, group = treatment), width = 0.2) + 
  geom_point(aes(y = mean_drinks), size = 2) +
  ggthemes::theme_few() + ggthemes::scale_color_few()+ labs(title = "Mean, Trajectories by Treatment", y = "Drinks", x = "Days")


alc_dat_long %>% 
  group_by(treatment, days) %>% 
  summarize(mean_drinks = mean(drinks),
            upper = mean_drinks + sd(drinks),
            lower = mean_drinks - sd(drinks)) %>% 
  ggplot(aes(x = days, color = treatment)) +
  geom_line(aes(y = mean_drinks, group = treatment), size = 1) + 
  geom_errorbar(aes(ymin = lower, ymax = upper, group = treatment), width = 0.2) + 
  geom_point(aes(y = mean_drinks), size = 2) +
  geom_line(data = alc_dat_long, aes(x = days, y = drinks, group = sid), alpha = 0.2) + 
  ggthemes::theme_few() + ggthemes::scale_color_few() + facet_grid(~treatment) + labs(title = "Raw Trajectories", y = "Drinks", x = "Days")


```

There are clearly two groups of patients in all groups, with a wide berth between the groups: those whose baseline number of alcoholic drinks in the past 30 days is less than  100 (~3 drinks/day) and those who are closer to 150 in the past 30 days (5 drinks a day).

Perhaps this could be a gender effect:

```{r}
alc_dat_long %>% 
  group_by(treatment, days, gender) %>% 
  summarize(mean_drinks = mean(drinks),
            upper = mean_drinks + sd(drinks),
            lower = mean_drinks - sd(drinks)) %>% 
  ggplot(aes(x = days, color = treatment)) +
  geom_line(data = alc_dat_long, aes(x = days, y = drinks, group = sid), alpha = 0.2) +
  geom_line(aes(y = mean_drinks, group = treatment)) + 
  geom_errorbar(aes(ymin = lower, ymax = upper, group = treatment), width = 0.2) + 
  geom_point(aes(y = mean_drinks)) +
  ggthemes::theme_few() + ggthemes::scale_color_few() + facet_grid(~gender) + labs(title = "Raw Trajectories", y = "Drinks", x = "Days")
```

It is, in fact, a gender effect, and variance shrinks a great deal when the outcome is stratified on gender. Luckily, there does not seem to be heterogeneity of treatment effect by gender; the treatment trajectories look the same, but scaled to the respective within-gender mean.

Now, lets's look at faceting on relapsers vs. no relapsers:

```{r}
library(janitor)
alc_dat %>% 
  tabyl(treatment, relapse) %>% 
  adorn_percentages("row") %>%
  adorn_pct_formatting(digits = 2) %>%
  adorn_ns() %>% 
  knitr::kable(caption = "Contingency table of Relapse by Treatment")

alc_dat %>% 
  tabyl(treatment, relapse) %>% 
  chisq.test()
#significant differences. pairwise:

alc_dat %>% 
  filter(treatment %in% 1:2) %>% 
  mutate(treatment = forcats::fct_drop(treatment)) %>% 
  tabyl(treatment, relapse) %>% 
  chisq.test(.)

alc_dat %>% 
  filter(treatment %in% c(1,3)) %>% 
  mutate(treatment = forcats::fct_drop(treatment)) %>% 
  tabyl(treatment, relapse) %>% 
  chisq.test(.)

alc_dat %>% 
  filter(treatment %in% 2:3) %>% 
  mutate(treatment = forcats::fct_drop(treatment)) %>% 
  tabyl(treatment, relapse) %>% 
  chisq.test(.)



#all pairwise differences are highly significant

alc_dat_grouped <- alc_dat %>% 
  mutate(treatment = forcats::fct_relevel(treatment, "3")) %>% 
  group_by(treatment, gender) %>% 
  count(relapse) %>% 
  pivot_wider(names_from = relapse, values_from = n, names_prefix = "relapse_")

response_table = cbind(alc_dat_grouped$relapse_1, alc_dat_grouped$relapse_0)
treatment <- alc_dat_grouped$treatment
glm_relapse_ref <- glm(response_table ~ treatment , family = binomial) 

broom::tidy(glm_relapse_ref, exponentiate = TRUE, conf.int = TRUE) %>% 
    dplyr::select(term, estimate, std.error, conf.low, conf.high) %>% 
  knitr::kable(caption = "Logistic Regression of Treatment on 6-month Relapse")


broom::glance(glm_relapse_ref) %>% 
  summarize(test = 1 - pchisq(deviance, df.residual)) #acceptable

anova(glm_relapse_ref, test = "Chisq") #significant
```


# GEE models

## Exploratory Correlation Plots

```{r}
sat_mod <- glm(drinks ~ gender + treatment*days, family = "poisson", data = alc_dat_long) #super saturated model
alc_dat_long %>% 
  modelr::add_residuals(sat_mod) %>% 
  dplyr::select(sid, days, treatment, gender, resid) %>% 
  mutate(days = as.numeric(days)) %>% 
  group_by(sid) %>% 
  mutate(lag_1 = lag(resid, 1),
         lag_2 = lag(resid, 2)) %>% 
  ungroup() %>% group_by(gender, treatment) %>% 
  summarize(cor_0 = cor(resid, resid, use = "pairwise.complete.obs"),
            cor_1 = cor(resid, lag_1, use = "pairwise.complete.obs"),
            cor_2 = cor(resid, lag_2, use = "pairwise.complete.obs")) %>% ungroup() %>% 
  summarize(cor_0 = mean(cor_0), cor_1 = mean(cor_1), cor_2 = mean(cor_2)) %>% 
  pivot_longer(c("cor_0", "cor_1", "cor_2"), names_to = "lag", values_to = "cor", names_prefix = "cor_") %>%
  ggplot(aes(x = lag, y = cor)) +
  geom_col(width = 0.1) + 
  labs(title = "ACF with lagged measurements") 
```


# Fit GEE models

```{r}
pois_gee <- geeglm(drinks ~ treatment*days + gender,
       family = poisson,
       data = alc_dat_long,
       id = sid,
       corstr = "exchangeable")

summary(pois_gee)


pois_gee_int <- geeglm(drinks ~ treatment*days + treatment*gender,
       family = poisson,
       data = alc_dat_long,
       id = sid,
       corstr = "ar1")

summary(pois_gee_int) #ns

pois_gee_intd <- geeglm(drinks ~ treatment*days + days*gender,
       family = poisson,
       data = alc_dat_long,
       id = sid,
       corstr = "ar1")


broom::tidy(pois_gee_intd, exponentiate = TRUE, conf.int = TRUE) #significant effect for day 60 but very small

anova(pois_gee, pois_gee_int)
anova(pois_gee, pois_gee_intd)




##Trajectories##
#get contrast of treatment 3 days 60 to treatment 3 days 30
#get contrast of treatment 1 days 60 to treatment 1 days 30
#get contrast of treatment 3 days 60 to treatment 3 days 30

## Treatment effect ##
#get contrast of treatment 3 days 60 to treatment 2 days 60
#get contrast of treatment 3 days 60 to treatment 2 days 20

colnames(pois_gee$geese$vbeta) = names(coef(pois_gee))
rownames(pois_gee$geese$vbeta) = names(coef(pois_gee))
var_b0 = pois_gee$geese$vbeta["(Intercept)","(Intercept)"]
var_b1 = pois_gee$geese$vbeta["treatment2","treatment2"]
var_b2 = pois_gee$geese$vbeta["(Intercept)","(Intercept)"]
var_b3 = pois_gee$geese$vbeta["(Intercept)","(Intercept)"]
var_b4 = pois_gee$geese$vbeta["(Intercept)","(Intercept)"]

var_b6 = pois_gee$geese$vbeta["genderFemale","genderFemale"]
cov_b0b6 = pois_gee$geese$vbeta["genderFemale","(Intercept)"]
women_count_baseline = exp(coef(pois_gee)["(Intercept)"] + coef(pois_gee)["genderFemale"])
women_var_baseline = var_b0 + var_b6 + 2*cov_b0b6
women_ci_baseline = c(exp(coef(pois_gee)["(Intercept)"] + coef(pois_gee)["genderFemale"] - qchisq(0.975, 1)*sqrt(women_var_baseline)), exp(coef(pois_gee)["(Intercept)"] + coef(pois_gee)["genderFemale"] + qchisq(0.975, 1)*sqrt(women_var_baseline)))

var_b4 = pois_gee$geese$vbeta["days30","days30"]
var_b5 = pois_gee$geese$vbeta["days60","days60"]
cov_b4b5 = pois_gee$geese$vbeta["days30","days60"]
d30_to_d60_rr = exp(coef(pois_gee)["days60"] - coef(pois_gee)["days30"])
d30_d60_var = var_b4 + var_b5 - 2*cov_b4b5
d30_to_d60_ci = c(exp(coef(pois_gee)["days60"] - coef(pois_gee)["days30"] - qchisq(0.975, 1)*sqrt(d30_d60_var)), exp(coef(pois_gee)["days60"] - coef(pois_gee)["days30"] + qchisq(0.975, 1)*sqrt(d30_d60_var)))

#treatment 2 to treatment 1
var_trt2 = pois_gee$geese$vbeta["treatment2","treatment2"]
var_trt2d30 = pois_gee$geese$vbeta["treatment2:days30","treatment2:days30"]
cov_day30trt2 = pois_gee$geese$vbeta["treatment2:days30","treatment2"]
trt2day30totrt1day30_rr = exp(coef(pois_gee)["treatment2"] + coef(pois_gee)["treatment2:days30"])
trt2day30totrt1day30_var = var_trt2 + var_trt2d30 - 2*cov_day30trt2
trt2day30totrt1day30_ci = c(exp(log(trt2day30totrt1day30_rr) - qchisq(0.975, 1)*sqrt(trt2day30totrt1day30_var)), exp(log(trt2day30totrt1day30_rr) + qchisq(0.975, 1)*sqrt(trt2day30totrt1day30_var)))

var_trt2d60 = pois_gee$geese$vbeta["treatment2:days60","treatment2:days60"]
cov_day60trt2 = pois_gee$geese$vbeta["treatment2:days60","treatment2"]
trt2day60totrt1day60_rr = exp(coef(pois_gee)["treatment2"] + coef(pois_gee)["treatment2:days60"])
trt2day60totrt1day60_var = var_trt2 + var_trt2d60 + 2*cov_day60trt2
trt2day60totrt1day60_ci = c(exp(log(trt2day60totrt1day60_rr) - qchisq(0.975, 1)*sqrt(trt2day60totrt1day60_var)), exp(log(trt2day60totrt1day60_rr) + qchisq(0.975, 1)*sqrt(trt2day60totrt1day60_var)))

cov_day60day30trt2 = pois_gee$geese$vbeta["treatment2:days60","treatment2:days30"]
trt2trt1_day60today30_r_rr = exp(coef(pois_gee)["treatment2:days60"] - coef(pois_gee)["treatment2:days30"])
trt2trt1_day60today30_r_var = var_trt2d30 + var_trt2d60 - 2*cov_day60day30trt2
trt2trt1_day60today30_r_ci = c(exp(log(trt2trt1_day60today30_r_rr) - qchisq(0.975, 1)*sqrt(trt2trt1_day60today30_r_var)), exp(log(trt2trt1_day60today30_r_rr) + qchisq(0.975, 1)*sqrt(trt2trt1_day60today30_r_var)))

#treatment 3 to treatment 1
var_trt3 = pois_gee$geese$vbeta["treatment3","treatment3"]
var_trt3d30 = pois_gee$geese$vbeta["treatment3:days30","treatment3:days30"]
cov_day30trt3 = pois_gee$geese$vbeta["treatment3:days30","treatment3"]
trt3day30totrt1day30_rr = exp(coef(pois_gee)["treatment3"] + coef(pois_gee)["treatment3:days30"])
trt3day30totrt1day30_var = var_trt3 + var_trt3d30 - 2*cov_day30trt3
trt3day30totrt1day30_ci = c(exp(log(trt3day30totrt1day30_rr) - qchisq(0.975, 1)*sqrt(trt3day30totrt1day30_var)), exp(log(trt3day30totrt1day30_rr) + qchisq(0.975, 1)*sqrt(trt3day30totrt1day30_var)))

var_trt3d60 = pois_gee$geese$vbeta["treatment3:days60","treatment3:days60"]
cov_day60trt3 = pois_gee$geese$vbeta["treatment3:days60","treatment3"]
trt3day60totrt1day60_rr = exp(coef(pois_gee)["treatment3"] + coef(pois_gee)["treatment3:days60"])
trt3day60totrt1day60_var = var_trt3 + var_trt3d60 + 2*cov_day60trt3
trt3day60totrt1day60_ci = c(exp(log(trt3day60totrt1day60_rr) - qchisq(0.975, 1)*sqrt(trt3day60totrt1day60_var)), exp(log(trt3day60totrt1day60_rr) + qchisq(0.975, 1)*sqrt(trt3day60totrt1day60_var)))

cov_day60day30trt3 = pois_gee$geese$vbeta["treatment3:days60","treatment3:days30"]
trt3trt1_day60today30_r_rr = exp(coef(pois_gee)["treatment3:days60"] - coef(pois_gee)["treatment3:days30"])
trt3trt1_day60today30_r_var = var_trt3d30 + var_trt3d60 - 2*cov_day60day30trt3
trt3trt1_day60today30_r_ci = c(exp(log(trt3trt1_day60today30_r_rr) - qchisq(0.975, 1)*sqrt(trt3trt1_day60today30_r_var)), exp(log(trt3trt1_day60today30_r_rr) + qchisq(0.975, 1)*sqrt(trt3trt1_day60today30_r_var)))


#treatment 3 to treatment 2
cov_day30trt3trt2 = pois_gee$geese$vbeta["treatment3:days30","treatment2"]
cov_day30trt3day30trt2 = pois_gee$geese$vbeta["treatment3:days30","treatment2:days30"]
cov_trt3day30trt2 = pois_gee$geese$vbeta["treatment3","treatment2:days30"]
cov_trt3trt2 = pois_gee$geese$vbeta["treatment3","treatment2"]
trt3day30totrt2day30_rr = exp(coef(pois_gee)["treatment3"] + coef(pois_gee)["treatment3:days30"] - 
                                coef(pois_gee)["treatment2"] - coef(pois_gee)["treatment2:days30"])
trt3day30totrt2day30_var = var_trt3 + var_trt3d30 + var_trt2 + var_trt2d30 + 2*cov_day30trt3 + 2*cov_day30trt2 - 2*cov_day30trt3trt2 -2*cov_day30trt3day30trt2 -2*cov_trt3day30trt2 - 2*cov_trt3trt2
trt3day30totrt2day30_ci = c(exp(log(trt3day30totrt2day30_rr) - qchisq(0.975, 1)*sqrt(trt3day30totrt2day30_var)), exp(log(trt3day30totrt2day30_rr) + qchisq(0.975, 1)*sqrt(trt3day30totrt2day30_var)))

cov_day60trt3trt2 = pois_gee$geese$vbeta["treatment3:days60","treatment2"]
cov_day60trt3day60trt2 = pois_gee$geese$vbeta["treatment3:days60","treatment2:days60"]
cov_trt3day60trt2 = pois_gee$geese$vbeta["treatment3","treatment2:days60"]
cov_trt3trt2 = pois_gee$geese$vbeta["treatment3","treatment2"]
trt3day60totrt2day60_rr = exp(coef(pois_gee)["treatment3"] + coef(pois_gee)["treatment3:days60"] - 
                                coef(pois_gee)["treatment2"] - coef(pois_gee)["treatment2:days60"])
trt3day60totrt2day60_var = var_trt3 + var_trt3d30 + var_trt2 + var_trt2d30 + 2*cov_day30trt3 + 2*cov_day30trt2 - 2*cov_day60trt3trt2 -2*cov_day30trt3day30trt2 -2*cov_trt3day30trt2 - 2*cov_trt3trt2
trt3day60totrt2day60_ci = c(exp(log(trt3day60totrt2day60_rr) - qchisq(0.975, 1)*sqrt(trt3day60totrt2day60_var)), exp(log(trt3day60totrt2day60_rr) + qchisq(0.975, 1)*sqrt(trt3day60totrt2day60_var)))



broom::tidy(pois_gee, exponentiate = TRUE, conf.int = TRUE) %>% 
  dplyr::select(term, estimate, std.error, conf.low, conf.high) %>% 
  knitr::kable(caption = "Ratio of expected counts from Poisson GEE model (Exchangeable covariance)")

```



